---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bedoc

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![R build status](https://github.com/tjtnew/bedoc/workflows/R-CMD-check/badge.svg)](https://github.com/tjtnew/bedoc/actions)
[![CRAN status](https://www.r-pkg.org/badges/version/bedoc)](https://CRAN.R-project.org/package=bedoc)
<!-- badges: end -->

> Hospital Bed Occupancy Forecasting

`bedoc` implements forecasting of daily bed occupancy from input data on daily
admissions and the distribution of duration of stay.  It aims to provide a
reliable and standardised approach to forecasting that will help improve
the quality and ease of provision of predictions going forward.

## Installing the package
Whilst this package is not yet on cran the development, *github* version can be installed using:

```{r install2, eval = FALSE}
devtools::install_github("tjtnew/bedoc")
```

Note that this requires the package *devtools* to be installed.

## Functionality

The user is provided with a function `project_beds` that ... 

tbc


