---
title: "Overview of the bedoc package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(bedoc)
library(ggplot2)  # for plotting
library(scales)   # for plotting
```

# Introduction
`bedoc` implements forecasting of daily bed occupancy from input data on daily
admissions and the distribution of duration of stay.  It aims to provide a
reliable and standardised approach to forecasting that will help improve
the quality and ease of provision of predictions going forward.

## Available functions
`bedoc` contains just four functions, three of which are reexported from other
packages for user convenience:

* `project_beds` is the main function to compute projections of bed occupancy
from admissions data and a length of stay distribution.
* `build_projections` is reexported from the 
[`projections`](https://cran.r-project.org/web/packages/projections/index.html)
package.  It allows the user to build a
\code{\link[projections:build_projections]{projections}}
object that can be fed to the `project_beds` function.
* `incidence` is reexported from the 
[`incidence`](https://cran.r-project.org/web/packages/incidence/index.html)
package.  Similarly to `build_projections` function above, it allows the user to
build a \code{\link[incidence::incidence]{incidence}} object that can be fed to
`project_beds`.

# Usage
We consider two examples where we have hospital admission data and our goal is
to predict bed occupancy for the following weeks. With both examples we will 
consider an outbreak in it's early stages where we are anticipating exponential
growth in the number of cases.

Both examples we will use the
[Weibull distribution](https://en.wikipedia.org/wiki/Weibull_distribution) for a patient's length of stay and we use the
[`distcrete`](https://cran.r-project.org/web/packages/distcrete/index.html)
package to discretise the distribution.


```{r}
library(distcrete)
length_of_stay <-
    distcrete("weibull", shape = 1.1, scale = 7.4, w = 0.5, interval = 1)
```


## Example 1
In this example we assume the disease has a doubling time of 7 days and 
that there are initially 10 hospital admissions.

```{r}
initial_admissions <- 10 # day one
duration <- 14           # 14 days (2 weeks) duration
growth_rate = log(2) / 7 # doubling time of 7 days

future_admissions <- round(
    initial_admissions * exp(growth_rate * (seq_len(duration))))

admissions <- 
  incidence(rep(Sys.Date() + 0:14, c(initial_admissions, future_admissions)))

# convert to a projections object to use with the project_beds function
admissions <- build_projections(
    x = c(initial_admissions, future_admissions), 
    dates = Sys.Date() + 0:14)
```

We now have all the variables we need to make our forecast of bed occupancy

```{r}
projection <- project_beds(x = admissions, 
                           r_los = length_of_stay, 
                           n_sims = 10)

plot(projection, quantiles = c(0.025, 0.5)) + 
    scale_x_date(breaks = breaks_pretty(10)) + 
    ylim(0, 500)
```

## Example 2
In this example we assume we have a weeks worth of admission data that, again, 
we want to forecast the next 14 days.  This time, however, we would like to
encorporate the disease serial-interval and a value of R0.

To start with we make up some toy admissions data, use a value of R0 = 2.5 and
create a discretised serial interval.  To give the example some some grounding 
in reality we use some reasearch from the recent COVID-19 pandemic.  In a small [study](https://www.ijidonline.com/article/S1201-9712(20)30119-3/fulltext) it
was observed that a log-normal distribution provided a good fit for
data that had an observed mean of 4.7 days and standard deviation of 2.9 days.

```{r}
library(projections)  # so we can use the `project` and subset functions 

# make some fake admissions data
initial_admissions <- sample(Sys.Date() - 0:7, 30, replace = TRUE)
initial_admissions <- incidence(initial_admissions)

# observed study values need converting before use with R's log-normal function
recorded_mean <- 4.7
recorded_sd <- 2.9
mu <- log(recorded_mean^2 / sqrt(recorded_mean^2 + recorded_sd^2))
sd = sqrt(log(1 + recorded_sd^2 / recorded_mean^2))

# generate the discretised serial inteval
serial_interval <- distcrete("lnorm", mu, sd, w = 0.5, interval = 1)
plot(1:20, serial_interval$d(1:20), type = "h", lwd = 5, col = "navy",
     main = "Serial interval", xlab = "Days after onset",
     ylab = "Relative infectiousness")

# project future admissions
future_admissions <- project(initial_admissions,
                             R = c(2,2.5), si = serial_interval,
                             n_days = 14, n_sim = 100)

# combine the current and future admissions data
original_admissions <- build_projections(initial_admissions$counts,
                                        initial_admissions$dates)
admissions <- future_admissions + original_admissions
```

Now we can again make our forecasts
```{r}
projections <- project_beds(x = admissions, 
                           r_los = length_of_stay, 
                           n_sims = 10)

projections <- subset(projections, from = Sys.Date())

plot(projections, quantiles = c(0.025, 0.5)) + 
    scale_x_date(breaks = breaks_pretty(10)) + 
    ylim(0, 500)
```
